Notes

### 2022-09-19 chapters/experiments.tex
Small changes after revision

Because of execution time constraints, the hyperparameters of the MCMC method are not part of the CV process, and are selected manually based on an initial set of experiments, as well as recommendations from the original article. We use 64 chains and run them for 900 iterations in total, discarding the first 500 iterations as burn-in. Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains at each iteration, selecting the stretch move (the default) with probability 0.7 or the walk move with probability 0.3.

@@

\paragraph{Linear regression data sets.} We consider three different types of simulated data sets, all with a common value of \(\alpha_0=5\).
\begin{itemize}
  \item A RKHS model where the response is generated as \(Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
  \item A \(L^2\)-model with underlying coefficient function \(\beta(t)=\log(1+4t)\), again for the same four GP's.
  \item A model based on non-GP regressors. Specifically, we use a geometric Brownian motion (GBM) defined as \(X(t)=e^{BM(t)}\), where \(BM(t)\) is a standard Brownian motion. In this case, we consider two data sets, one with a RKHS response and one with a \(L^2\) response, with the same parameters as above.
\end{itemize}
As for the real data sets, we use the (twice-differentiated) Tecator data set introduced in Chapter~\ref{ch:introduction} to predict fat content of 193 meat samples, as well as what we call the Moisture \citep{kalivas1997two} and Sugar \citep{bro1999exploratory} data sets. The first consists of near-infrared spectra of 100 wheat samples and the objective is to predict the samples' moisture content, whereas the second contains 268 samples of sugar fluorescence data in order to predict ash content. The three data sets are measured on a grid of 100, 101 and 115 equispaced points on \([0, 1]\), respectively.

\paragraph{Logistic regression data sets.} Again we consider three different types of simulated data sets,  with a common value of \(\alpha_0=-0.5\). In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
\begin{itemize}
  \item Four logistic RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
  \[
  \P(Y=1\mid X) = \frac{1}{1 + \exp\left\{0.5 +5X(0.1) - X(0.4) - 10X(0.8)\right\}}.
  \]
  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case, i.e.:
  \[
  \P(Y=1\mid X) = \frac{1}{\displaystyle 1 + \exp\left\{0.5 - \int_0^1 \log(1+4t) X(t)\, dt\right\}}.
  \]
  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's (with equal probability \(p=1/2\)) and label them according to their origin. Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\). Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance 2, that is, with kernel \(K(t,s)=2\min\{t,s\}\).
\end{itemize}
In this case we use three real data sets well known in the literature. The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, trying to predict if their longevity is high or low. The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives. Finally, we have selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.

@@

\paragraph{Linear regression comparison algorithms.} We consider the following regression methods:
\begin{itemize}
  \item Manual.\hspace{.3em} Dummy variable selection method with a pre-specified number of components (equispaced on \([0, 1]\)).
  \item Lasso.\hspace{.3em} Linear least squares with \(l^1\) regularization.
  \item Ridge.\hspace{.3em} Linear least squares with \(l^2\) regularization.
  \item PLS.\hspace{.3em} Partial least squares for dimension reduction.
  \item PCA.\hspace{.3em} Principal component analysis for dimension reduction.
  \item PLS1.\hspace{.3em} Partial least squares regression \citep[e.g.][]{wegelin2000survey}.
  \item APLS.\hspace{.3em} Functional partial least squares regression proposed by \citet{delaigle2012methodology}.
  \item RMH.\hspace{.3em} Recursive maxima hunting variable selection method proposed by \citet{torrecilla2016feature}.
  \item FLin.\hspace{.3em} Functional \(L^2\) linear regression model with fixed basis expansion.
  \item FPLS1.\hspace{.3em} Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
  \item FPCA.\hspace{.3em} Functional principal component analysis.
\end{itemize}

@@

The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair. In particular, the regularization parameter is searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for quadratic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).

Most algorithms have been taken from the libraries \textit{scikit-learn}\footnote{\url{https://scikit-learn.org/stable/}} and \textit{scikit-fda}\footnote{\url{https://fda.readthedocs.io/en/latest/}}, the first oriented to machine learning in general and the second to functional data analysis in particular. However, some methods were not found in these packages and had to be implemented from scratch. This is the case of the FLDA, FPLS and APLS methods, which we coded following the corresponding articles.

@@

In Figure~\ref{fig:reg_emcee_rkhs} we see the results for the four GP regressors considered in the RKHS case. This is the most favorable case for us, as the underlying model coincides with our assumed model. Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably with the Gaussian kernel. A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more customized approach would be needed to obtain better results.

@@

Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic processes, but they struggle in the heteroscedastic case. Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e. with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}. Our results are in line with the empirical studies of said article, where the authors conclude that even though the asymptotic theoretical error is zero, most classification methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.

@@

In this section we present some experiments that are not directly related to measuring the prediction performance of our models relative to other methods.

@@

One requirement that our model should satisfy is that it should be able to recover the true parameter function when the underlying data generation model is a finite-dimensional RKHS model. This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?

Take for example a RKHS data set with two components generated according to the formula \(Y=5 -5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\). Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(\tau_1, \tau_2, \tau_3)\) for a model with 3 components. As we can see, one of the coefficients has gone to zero to account for the overspecification of the model, while the other two have stabilized very close to the true parameters. The same goes for the time instants, except that in this case there is no default value to represent that a component is unused, so the time corresponding to the null coefficient oscillates back and forth. The estimated function (based for example in the mode of the posterior distributions) will not be perfect, essentially because of the noise in the response. But it should be close to the true parameter function \(\alpha(t)=-5K(t, 0.1) + 10K(t, 0.8)\), providing a good predictive performance.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_3}
  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_4}
  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 700 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
\end{figure}

@@

Another thing we wanted to look at was the dependence of the final prediction result on the chosen value of \(p\), especially when there is no concept of ``components'' in the underlying model. We can take for example the homoscedastic mixture data set described earlier for the logistic regression problem, and fix the parameter \(\eta=0.01\). Then the corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).

@@

We also tried changing some prior distributions. For instance, we swapped the improper prior distribution on \(\sigma^2\) for a Cauchy distribution on \(\R^+\). The results were similar, but we ultimately discarded this approach because it introduced yet another hyperparameter into the model with no clear benefit. In addition, we tried to impose a beta distribution on each of the time instants \(\{t_j\}\), selecting the corresponding shape parameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable). This approach does work, but in our estimation the computational overhead introduced by the beta density function (sometimes resulting in a 1.5x increase in execution time) is not worth the mild improvements, if any.

>>>>

Because of execution time constraints, the hyperparameters of the MCMC method are not part of the CV process, and are selected manually based on an initial set of experiments, as well as recommendations from the original article. We use 64 chains and run them for 900 iterations in total, discarding the first 500 iterations as burn-in. Moreover, we use a weighted mixture of walk and stretch moves in the \textit{emcee} sampler to advance the chains in each iteration, selecting the stretch move (the default) with probability 0.7 or the walk move with probability 0.3.

@@

\paragraph{Linear regression data sets.} We consider three different types of simulated data sets, all with a common value of \(\alpha_0=5\).
\begin{itemize}
  \item A finite-dimensional RKHS model where the response is generated as \(Y=5 -5X(0.1) + X(0.4) + 10X(0.8) + \epsilon\), for each of the four GP's mentioned above.
  \item A \(L^2\)-model with underlying coefficient function \(\beta(t)=\log(1+4t)\), again for the same four GP's.
  \item A model based on non-GP regressors. Specifically, we use a geometric Brownian motion (GBM) defined as \(X(t)=e^{BM(t)}\), where \(BM(t)\) is a standard Brownian motion. In this case we consider two data sets, one with a RKHS response and one with a \(L^2\) response, with the same parameters as above.
\end{itemize}
As for the real data sets, we use the (twice-differentiated) Tecator data set introduced in Chapter~\ref{ch:introduction} to predict fat content of 193 meat samples, as well as what we call the Moisture \citep{kalivas1997two} and Sugar \citep{bro1999exploratory} data sets. The first consists of near-infrared spectra of 100 wheat samples and the objective is to predict the samples' moisture content, whereas the second contains 268 samples of sugar fluorescence data in order to predict ash content. The three data sets are measured on a grid of 100, 101 and 115 equispaced points on \([0, 1]\), respectively.

\paragraph{Logistic regression data sets.} Again we consider three different types of simulated data sets,  with a common value of \(\alpha_0=-0.5\). In this case we randomly permute 10\% of the labels to introduce some noise in the simulations.
\begin{itemize}
  \item Four logistic finite-dimensional RKHS models with the same functional parameter as in the linear regression case (one for each GP). Specifically,
  \[
  \P(Y=1\mid X) = \frac{1}{1 + \exp\left\{0.5 +5X(0.1) - X(0.4) - 10X(0.8)\right\}}.
  \]
  \item Four logistic \(L^2\)-models with the same coefficient function as in the linear regression case, i.e.:
  \[
  \P(Y=1\mid X) = \frac{1}{\displaystyle 1 + \exp\left\{0.5 - \int_0^1 \log(1+4t) X(t)\, dt\right\}}.
  \]
  \item A ``mixture'' model based on non-GP regressors, in which we mix regressors from two different GP's (with equal probability \(p=1/2\)) and label them according to their origin. Firstly, we consider a homoscedastic case to distinguish between a standard Brownian motion and a Brownian motion with a mean function that is zero until \(t=0.5\), and then becomes \(m(t)=0.75t\). Secondly, we consider a heteroscedastic case to distinguish between a standard Brownian motion and a Brownian motion with variance 2, that is, with kernel \(K(t,s)=2\min\{t,s\}\).
\end{itemize}
Additionally, we use three real data sets well known in the literature. The first one is a subset of the Medflies data set \citep{carey1998relationship}, consisting on samples of the number of eggs laid daily by 534 flies over 30 days, to predict if their longevity is high or low. The second one is the Berkeley Growth Study data set \citep{tuddenham1954physical}, which records the height of 54 girls and 39 boys over 31 different points in their lives. Finally, we selected a subset of the Phoneme data set \citep{hastie1995penalized}, based on 200 digitized speech frames over 128 equispaced points to predict the phonemes ``aa'' and ``ao''.

@@

\paragraph{Linear regression comparison algorithms.} We consider the following regression methods:
\begin{itemize}
  \item Manual.\hspace{.3em} Dummy variable selection method with a pre-specified number of components (equispaced on \([0, 1]\)).
  \item Lasso.\hspace{.3em} Linear least squares with \(l^1\) regularization.
  \item Ridge.\hspace{.3em} Linear least squares with \(l^2\) regularization.
  \item PLS.\hspace{.3em} Partial least squares for dimension reduction.
  \item PCA.\hspace{.3em} Principal component analysis for dimension reduction.
  \item PLS1.\hspace{.3em} Partial least squares regression \citep[e.g.][]{wegelin2000survey}.
  \item APLS.\hspace{.3em} Functional partial least squares regression proposed by \citet{delaigle2012methodology}.
  \item RMH.\hspace{.3em} Recursive maxima hunting variable selection method proposed by \citet{torrecilla2016feature}.
  \item FLin.\hspace{.3em} Functional \(L^2\) linear regression model with fixed basis expansion.
  \item FPCA.\hspace{.3em} Functional principal component analysis.
  \item FPLS1.\hspace{.3em} Functional PLS regression through basis expansion, implemented as in \citet{aguilera2010using}.
\end{itemize}

@@

The main parameters of all these algorithms are selected by cross-validation, using the same 5 folds as our proposed models so that the comparisons are fair. In particular, regularization parameters are searched among 20 values in the logarithmic space \([10^{-4}, 10^4]\), the number of manually selected variables is one of \(\{5, 10, 15, 20, 25, 50\}\), the number of components for dimension reduction and variable selection techniques is in the set \(\{2, 3, 4, 5, 7, 10, 15, 20\}\), the number of basis elements for cubic spline bases is in \(\{8,10,12,14,16\}\), the number of basis elements for Fourier bases is one of \(\{3,5,7,9,11\}\), and the number of neighbors in the KNN classifier is in \(\{3,5,7,9,11\}\).

Most algorithms have been taken from the libraries \textit{scikit-learn}\footnote{\url{https://scikit-learn.org}} and \textit{scikit-fda}\footnote{\url{https://fda.readthedocs.io}}, the first oriented to machine learning in general and the second to functional data analysis in particular. However, some methods were not found in these packages and had to be implemented from scratch. This is the case of the FLDA, FPLS and APLS methods, which we coded following the corresponding articles.

@@

In Figure~\ref{fig:reg_emcee_rkhs} we see the results for the four GP regressors considered in the RKHS case. This is the most favorable case for us, as the underlying model coincides with our assumed model. Indeed, we can see that in most cases our algorithms are the ones with lower RMSE, save for a few exceptions, notably the Gaussian kernel. A subsequent analysis showed that this particular data set is especially sensitive to the value of the hyperparameter \(\eta\) in our prior distribution, so a more customized approach would be needed to obtain better results.

@@

Finally, Figure~\ref{fig:clf_emcee_nongp} shows that our classifiers perform better than most comparison algorithms when separating two homoscedastic Gaussian processes, but they struggle in the heteroscedastic case. Incidentally, this heteroscedastic case of two zero-mean Brownian motions has a special interest, since it can be shown that the Bayes error is zero in the limit of dense monitoring (i.e. with an arbitrarily fine measurement grid), a manifestation of the ``near-perfect'' classification phenomenon analyzed for example in \citet{torrecilla2020optimal}. Our results are in line with the empirical studies of this article, where the authors conclude that even though the asymptotic theoretical error is zero, most classification methods are suboptimal in practice (possibly due to the high collinearity of the data), with the notable exception of PCA+QDA.

@@

In this section we present some experiments that are not directly related to measuring the predictive performance of our models relative to other methods.

@@

One requirement that our model should satisfy is that it ought to be able to recover the true parameter function when the underlying data generation model is a finite-dimensional RKHS model. This is generally the case when the value of \(p\) in our model and the true value of \(p\) coincide, but what happens when we change the value of \(p\) in the model?

Take for example a RKHS data set with two components generated according to the formula \(Y=5 -5X(0.1) + 10X(0.8) + \epsilon\), with \(\epsilon\sim \mathcal N(0, 0.5)\). Figure~\ref{fig:beta_trace_3} shows the resulting posterior distribution of the parameters \(b=(\beta_1, \beta_2, \beta_3)\) and \(\tau=(t_1, t_2, t_3)\) for a model with 3 components. As we can see, one of the coefficients has gone to zero to account for the overspecification of the model, while the other two have stabilized very close to the true parameters. The same goes for the time instants, except that in this case there is no default value to represent that a component is unused, so the time corresponding to the null coefficient oscillates back and forth. The estimated function (based for example in the mode of the posterior distributions) will not be perfect, essentially because of the noise in the response. But it should be close to the true parameter function \(\alpha(t)=-5K(t, 0.1) + 10K(t, 0.8)\), providing a good predictive performance.

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_3}
  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=3\) (left) and the corresponding trace evolution for 400 iterations in the MCMC sampler (right).}\label{fig:beta_trace_3}
\end{figure}

@@

\begin{figure}[ht!]
  \centering
  \includegraphics[width=.8\textwidth]{beta_trace_4}
  \caption{Estimated posterior distribution for \(b\) and \(\tau\) in a linear model with \(p=4\) (left) and the corresponding trace evolution for 400 iterations in the MCMC sampler (right).}\label{fig:beta_trace_4}
\end{figure}

@@

Another thing we wanted to look at was the dependence of the final prediction result on the chosen value of \(p\), especially when there is no concept of ``components'' in the underlying model. We can take for example the homoscedastic mixture data set described earlier for the logistic regression problem, and fix the parameter \(\eta=0.01\). The corresponding mean accuracies (in 10 repetitions) for RKHS models with \(p=1,2,\dots 10\) components are shown in Figure~\ref{fig:dependence_acc_p}, along with their standard errors (which are arguably not very informative).

@@

We also tried changing some prior distributions. For instance, we swapped the improper prior distribution on \(\sigma^2\) for a Cauchy distribution on \(\R^+\). The results were similar, but we ultimately discarded this approach because it introduced yet another hyperparameter into the model with no clear benefit. In addition, we tried to impose a beta distribution on each of the time instants \(t_j\), selecting the corresponding shape parameters in a semi-automatic way so that most of the density fell in the areas where the variance of the process was higher (and thus the effect of individual trajectories was more noticeable). This approach does work, but in our estimation the computational overhead introduced by the beta density function (sometimes resulting in a 1.5x increase in execution time) is not worth the mild improvements, if any.

### 2022-09-19 chapters/introduction.tex
Small changes after revision

In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations. Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions. Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios widely explored in the literature. To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e. a process with finite second moments and whose realizations are square-integrable functions indexed on \([0,1]\)). The underlying probability space \((\Omega, \mathcal A, \P)\) is not important. We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\). In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid \(\{t_j\}\subset [0, 1]\). Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\). Figure~\ref{fig:linear_data_example} depicts a typical data set used in functional regression, and we already saw in Figure~\ref{fig:tecator} what a functional classification data set may look like.

>>>>

In this work we are concerned with functional linear and logistic regression models, that is, situations where the goal is to predict a continuous or dichotomous variable from functional observations. Even though these problems can be formally stated with almost no differences from their finite-dimensional counterparts, there are some fundamental challenges as well as some subtle drawbacks that emerge as a result of working in infinite dimensions. Moreover, we will concentrate our efforts on the case in which the response is a scalar, though function-on-function and function-on-scalar regression are also interesting scenarios widely explored in the literature. To set a common framework, throughout this work we will consider a scalar response variable \(Y\) (either continuous or binary) which has some dependence on a stochastic \(L^2\)-process \(X=X(t)=X(t, \omega)\) with trajectories in \(L^2[0, 1]\) (i.e. a process with finite second moments and whose realizations are square-integrable functions indexed on \([0,1]\)). The underlying probability space \((\Omega, \mathcal A, \P)\) is not important. We will further suppose that \(X\) is centered, that is to say, its mean function \(m(t)=\E[X(t)]\) vanishes for all \(t\in[0,1]\). In addition, we will tacitly assume the existence of a \textit{labeled} data set \(\D_n =\{(X_i, Y_i): i=1,\dots, n\}\) of independent observations from \((X, Y)\), where the functional observations are recorded on a common finite grid on \([0, 1]\). Our ultimate aim will be to accurately predict the response corresponding to unlabeled samples from \(X\). Figure~\ref{fig:linear_data_example} depicts a typical data set used in functional regression, and we already saw in Figure~\ref{fig:tecator} what a functional classification data set may look like.

### 2022-09-19 chapters/model-choice.tex
Small changes after revision

The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein. A common approach that seems to work well is to simply impose an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with \(\tau\). We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).

This is an area of ongoing research, and thus there are other, more complex relabeling strategies, both deterministic and probabilistic. A summary of several such methods can be found for example in \citet{rodriguez2014label} and \citet{papastamoulis2015label}. In particular, we tested the pivot method proposed by \citet{marin2005bayesian}, in which all samples are aligned to minimize their distance to a reference element (the ``pivot''), which is chosen as the sample that maximizes the posterior density. However, the process of finding the appropriate permutation in each case was time-consuming, and the results were similar, if not worse, than the ones obtained with the simpler order constraints, so the latter were chosen as the default relabeling method of our algorithm (though the pivot method can still be enabled through a dedicated argument in the sampling procedure).

@@

Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. These methods are usually based in computing a quantity known as the \textit{Bayes factor}, which in turn needs the normalizing integral constant we have been trying to avoid all along. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, cross-validation, \ldots) to select its optimal value. As we will see shortly in Chapter~\ref{ch:experiments}, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios.

>>>>

The most straightforward solutions consist on imposing an artificial identifiability constraint on the parameters to break the symmetry of their posterior distributions; see \citet{jasra2005markov} and references therein. A common approach that seems to work well is to simply enforce an ordering in the parameters in question, which in our case would mean requiring for example that \(\beta_i < \beta_j\) for \(i < j\), or the analogous with the times \(\tau\). We have implemented a variation of this method described in \citet{simola2021approximate}, which works by post-processing the samples and relabeling the components to satisfy the order constraint mentioned above, choosing either \(b\) or \(\tau\) depending on which set of ordered parameters would produce the largest separation between any two of them (suitably averaged across all iterations of the chains).

This is an area of ongoing research, and thus there are other, more complex relabeling strategies, both deterministic and probabilistic. A summary of several such methods can be found for example in \citet{rodriguez2014label} and \citet{papastamoulis2015label}. In particular, we tested the pivot method proposed by \citet{marin2005bayesian}, in which all samples are aligned to minimize their distance to a reference element (the ``pivot''), which is chosen as the sample that maximizes the posterior density. However, the process of finding the appropriate permutation in each case was time-consuming, and the results were similar, if not worse, than the ones obtained with the simpler order constraints. For this reason, the latter were chosen as the default relabeling method of our algorithm, though the pivot method can still be enabled through a dedicated argument in the sampling procedure.

@@

Another possibility would be to adapt a purely Bayesian model selection technique to our framework \citep[see][]{piironen2017comparison, gelman2013bayesian}, or even derive some model aggregation methods to combine the posterior distributions obtained for different-sized models. These methods are usually based in computing a quantity known as the \textit{Bayes factor}, which in turn requires the specific value of the normalizing integral constant we have been trying to avoid all along. In the end, for the sake of simplicity we decided to let \(p\) be an hyperparameter, so that we could use any model selection criteria (e.g. BIC, DIC, cross-validation, \ldots) to select its optimal value. As we will see shortly in Chapter~\ref{ch:experiments}, the experiments carried out indicate that even low values of \(p\) provide sufficient flexibility in most scenarios, and they suggest that our models are not very sensitive to small variations of this parameter.

\enlargethispage{1\baselineskip}
